Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation

### 2013-11-18 writedown/npt-neq/alanine.tex
back up before revise according to Giovanni

\title{Exploring the conformational dynamics of a large molecule in solution using an external electric field: A nonequilibrium molecular dynamics simulation}
% \title{The non-equilibrium molecular dynamics simulation of the alanine dipeptide under electric field}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
\affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

\begin{abstract}
In this paper, we investigate the conformational dynamics of alanine
   dipeptide under an external electric field by 
   nonequilibrium molecular dynamics simulation.  We consider
   the case of a constant and of an oscillatory field. In this context we
propose a procedure to
   implement the temperature control, which removes the irrelevant thermal
effects of the field. For the constant field different time-scales are
   identified  in the conformational, dipole moment, and orientational
   dynamics. Moreover, we prove that the solvent structure only
   marginally changes when the external field is switched on.  In the
   case of oscillatory field, the conformational changes are shown to be as
   strong as the previous case, and non-trivial nonequilibrium
   circular paths in the conformation space are revealed by calculating the
   integrated net probability fluxes.

@@

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the dynamical approach to nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known. In particular, in our case,
it is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that he dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x, t)$, where $\mathcal H$ is the time-dependent Hamiltonian,
and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = U^\dagger(t,0) \rho(\vect x, 0)$,
where $U^\dagger(t,0) = \exp\{-i \int_0^t d t' L(t')\}$.
On the other hand,
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t)).
\end{align}
This equation can be formally solved by
$\hat O(\vect x(t)) = U(t,0)\,\hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), U^\dagger(t,0)\, \rho(\vect x, 0)\rangle
  = \langle U(t,0)\,\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle.
\end{align}
Since we assume that the system starts from the distribution of equilibrium (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable computed along trajectories, starting from
initial configurations sampled from an \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the full Hamiltonian dynamics is integrated until
time $t$.
In this paper, these trajectories are called \emph{branching
  trajectories} {\bf EITHER WE PUT THE PICTURE OR REMOVE THIS SENTENCE {(see Fig.\ref{branching} for a pictorial representation)}}. 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.
Stochastic dynamics (e.g.~Langevin dynamics) can be handled analogously.

@@

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and dissolved in 644 TIP3P
water molecules.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
First, an equilibrium NVT simulation at 300~K of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by GROMACS.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
\bluec{
  In all testing cases, this local thermostat is able to control
  the system at the desired temperature, i.e.~300~K (the results
  are not presented in this paper).
  \textit{NOTE: we should mention that the local thermostat works...}
}
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

\recheck{From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $C$; in fact it is \bluec{almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation}:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\, [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
\recheck{Where $\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to $\mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)$.}
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the timestep of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
A positive value \recheck{of $F_{J,I}(t)$} indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
\recheck{In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.}

@@

\redc{The dipole moment of the alanine dipeptide is defined by
\begin{align}\label{eqn:tmp8}
  \boldsymbol\mu_{\textrm{alanine}}(t) =
  \bigg\langle
  \sum_{i\in \{\textrm{atoms of alanine}\}}
  q_i\, \vect r_i(t)
  \bigg\rangle,
\end{align}
where $q_i$ denotes the partial charge of an atom $i$ on the alanine,
and the molecule is neutral, i.e.~$\sum_iq_i = 0$.
$\vect r_i$ denotes the position of the atom $i$.}
Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment 
reaches 85\% of its maximum value in only 20~ps, which is comparable to
the warm-up time $t_{warm}$, and is 25 times
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to its 
maximum value, that is, 
6.8~Debye, and  corresponds to an energy
\redc{($\boldsymbol\mu_{\textrm{alanine}} \cdot \vect E$)}
of c.a.~$-13.6$~kJ/mol.
We calculate the averaged dipole moment of different conformations
\redc{(defined by taking average in Eqn.~\eqref{eqn:tmp8}
only for a certain conformation)}
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
The dipole energy difference between 3.1 and 7.1~Debye is roughly 8.0~kJ/mol.
\recheck{Under a constant EF, the system will be likely to be driven towards those
conformations with higher dipole moment, because the
energy of the system will be lowered.  We suggest that this energy difference may be the reason why
we observe that the $\beta$-sheet conformations with lower dipole moment
are driven towards $\alpha$-helix conformations
$A_1$ and $C$, whose dipole moments are the highest
among all conformations. Moreover, the $B_2$ conformation vanishes because
its dipole is significantly lower than other conformations.}
\bluec{Comparing Fig.~\ref{fig:tmp3} with~\ref{fig:tmp5},
  the sharp increment of dipole moment before 20~ps is due to the quick vanishing of conformation
  $B_2$, while the slow increment of dipole moment until 400~ps is due to
  the slow migration from conformation $B_1$ to $A_1$ and $C$.
  \textit{NOTE: here we want to discuss the relation between
    the conformation and the dipole moment. I rewote this
    sentence.}
}
% From 0 to 20~ps, the probability of
% lowest dipole conformation $B_2$ quickly
% vanishes. We observe a sharp increment of the average molecular dipole and finally, approaching  500~ps, the conformation slowly migrates from
% $B_1$ to $A_1$ and $C$. Moreover, since the dipole moment $A_1$ and $C$
% is higher than $B_1$, the average molecular dipole moment slowly increases
% to the nonequilibrium steady value.

\subsection{Response of the orientational order parameter}
\recheck{As a complementary information to the behaviour of the molecular dipole moment it is of interest to describe the overall orientational behaviour of the molecules w.r.t. the EF. In fact while the dipole moment specifically expresses the rearrangement of charges within the molecule as a response to the EF, the overall direction of the molecule tells us about the interplay between the internal positional rearrangement of the atoms and their alignment w.r.t. the EF.   
To this aim,} let us define
the geometric direction of the alanine by the red vector in
Fig.~\ref{fig:tmp3}, i.e.~by the angle bisector of the two black vectors,
which connect the $\alpha$-carbon and the carbons
on the methyl groups. 
We considered the angle $\theta$ made by 
the orientation of the alanine dipeptide in space w.r.t.~the direction of EF
as a function of time.
Then we defined an orientational order parameter 
\begin{align}
  S_\theta = \langle 3\sin^2\theta - 2\rangle.
\end{align}
\recheck{The ensemble average is made along the trajectories, so it is
a time dependent observable.
The order parameter
indicates the (time dependent) average orientation of the molecule w.r.t. ~the direction of the EF.}
If the molecule is perfectly perpendicular to the EF, then $S_\theta = 1$, if is perfectly parallel then $S_\theta = -2$, if it has no directional
preference, then $S_\theta = 0$. 
From $t=0$ to roughly 10~ps, the orientational order parameter rapidly
decreases from 0 to $-0.15$, which means a weak alignment of the molecule to the external field.
Then from $t=0$ to 500~ps, the molecule slowly changes to the orientation
that is perpendicular to the EF.
\recheck{In fact,
  \bluec{the vector of the geometric direction tends to be perpendicular
    to the dipole moment,
    \textit{NOTE: $\frac{\pi}{2}$ corresponds to $S_\theta = 1$, however,
      we only have 0.4. So would should say a ``tendency of perpendicular''
      rather than ``perfectly perpendicular'' ($\frac{\pi}{2}$).}
  }  
  % the angle between the vector of the dipole moment and
  % the vector of the geometric direction is always close to
  % $\frac{\pi}{2}$;
  if one observes the correlation between the value
  of the dipole moment (which tends to be more parallel to the EF as
  its value increases) and the angle of the vector of the geometric
  direction, the result above can be easily explained.}
This orientational change is linked to the observed slowest
time scale of the molecular conformational change.

>>>>

\title{Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation}
% \title{The non-equilibrium molecular dynamics simulation of the alanine dipeptide under electric field}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
\affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

\begin{abstract}
In this paper, we investigate the conformational dynamics of alanine
   dipeptide under an external electric field by 
   nonequilibrium molecular dynamics simulation.  We consider
   the case of a constant and of an oscillatory field. In this context we
propose a procedure to
   implement the temperature control, which removes the irrelevant thermal
effects of the field. For the constant field different time-scales are
   identified  in the conformational, dipole moment, and orientational
   dynamics. Moreover, we prove that the solvent structure only
   marginally changes when the external field is switched on.  In the
   case of oscillatory field, the conformational changes are shown to be as
   strong as in the previous case, and non-trivial nonequilibrium
   circular paths in the conformation space are revealed by calculating the
   integrated net probability fluxes.

@@

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the dynamical approach to nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known. In particular, in our case,
it is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x, t)$, where $\mathcal H$ is the time-dependent Hamiltonian,
and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = U^\dagger(t,0) \rho(\vect x, 0)$,
where $U^\dagger(t,0) = \mathcal T\exp\{-i \int_0^t d t' L(t')\}$, and $\mathcal T$ is the time ordering operator.
On the other hand,
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t)).
\end{align}
This equation can be formally solved by
$\hat O(\vect x(t)) = U(t,0)\,\hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), U^\dagger(t,0)\, \rho(\vect x, 0)\rangle
  = \langle U(t,0)\,\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle.
\end{align}
Since we assume that the system starts from the equilibrium distribution (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable computed along trajectories, starting from
initial configurations sampled from an \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the full Hamiltonian dynamics is integrated until
time $t$.
In this paper, these trajectories are called \emph{branching
  trajectories}.
% {\bf EITHER WE PUT THE PICTURE OR REMOVE THIS SENTENCE {(see Fig.\ref{branching} for a pictorial representation)}}. 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.
Stochastic dynamics (e.g.~Langevin dynamics) can be handled analogously.

@@

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and dissolved in 644 TIP3P
water molecules.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
First, an equilibrium NVT simulation at 300~K of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by GROMACS.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
\recheck{
  In all testing cases, this local thermostat is able to control
  the system at the desired temperature, i.e.~300~K (the results
  are not presented in this paper).
}
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

\recheck{From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $C$; in fact it is {almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation}:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\, [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
\recheck{Where $\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to $\mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)$.}
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the timestep of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
A positive value \recheck{of $F_{J,I}(t)$} indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
\recheck{In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.}

@@

\redc{The dipole moment of the alanine dipeptide is defined by
\begin{align}\label{eqn:tmp8}
  \boldsymbol\mu_{\textrm{alanine}}(t) =
  \bigg\langle
  \sum_{i\in \{\textrm{atoms of alanine}\}}
  q_i\, \vect r_i(t)
  \bigg\rangle,
\end{align}
where $q_i$ denotes the partial charge of an atom $i$ on the alanine,
and the molecule is neutral, i.e.~$\sum_iq_i = 0$.
$\vect r_i$ denotes the position of the atom $i$.}
Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment 
reaches 85\% of its maximum value in only 20~ps, which is comparable to
the warm-up time $t_{warm}$, and is 25 times
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to its 
maximum value, that is, 
6.8~Debye, and  corresponds to an energy
\redc{($\boldsymbol\mu_{\textrm{alanine}} \cdot \vect E$)}
of c.a.~$-13.6$~kJ/mol.
We calculate the averaged dipole moment of different conformations
\redc{(defined by taking average in Eqn.~\eqref{eqn:tmp8}
only for a certain conformation)}
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
The dipole energy difference between 3.1 and 7.1~Debye is roughly 8.0~kJ/mol.
\recheck{Under a constant EF, the system will be likely to be driven towards those
conformations with higher dipole moment, because the
energy of the system will be lowered.  We suggest that this energy difference may be the reason why
we observe that the $\beta$-sheet conformations with lower dipole moment
are driven towards $\alpha$-helix conformations
$A_1$ and $C$, whose dipole moments are the highest
among all conformations. Moreover, the $B_2$ conformation vanishes because
its dipole is significantly lower than other conformations.}
\recheck{Comparing Fig.~\ref{fig:tmp3} with~\ref{fig:tmp5},
  the sharp increment of dipole moment before 20~ps is due to the quick vanishing of conformation
  $B_2$, while the slow increment of dipole moment until 400~ps is due to
  the slow migration from conformation $B_1$ to $A_1$ and $C$.
  % \textit{NOTE: here we want to discuss the relation between
  %   the conformation and the dipole moment. I rewote this
  %   sentence.}
}
% From 0 to 20~ps, the probability of
% lowest dipole conformation $B_2$ quickly
% vanishes. We observe a sharp increment of the average molecular dipole and finally, approaching  500~ps, the conformation slowly migrates from
% $B_1$ to $A_1$ and $C$. Moreover, since the dipole moment $A_1$ and $C$
% is higher than $B_1$, the average molecular dipole moment slowly increases
% to the nonequilibrium steady value.

\subsection{Response of the orientational order parameter}
\recheck{As a complementary information to the behaviour of the molecular dipole moment it is of interest to describe the overall orientational behaviour of the molecules w.r.t. the EF. In fact while the dipole moment specifically expresses the rearrangement of charges within the molecule as a response to the EF, the overall direction of the molecule tells us about the interplay between the internal positional rearrangement of the atoms and their alignment w.r.t. the EF.   
To this aim,} let us define
the geometric direction of the alanine by the red vector in
Fig.~\ref{fig:tmp3}, i.e.~by the angle bisector of the two black vectors,
which connect the $\alpha$-carbon and the carbons
on the methyl groups. 
We considered the angle $\theta$ made by 
the orientation of the alanine dipeptide in space w.r.t.~the direction of EF
as a function of time.
Then we defined an orientational order parameter 
\begin{align}
  S_\theta = \langle 3\sin^2\theta - 2\rangle.
\end{align}
\recheck{The ensemble average is made along the trajectories, so it is
a time dependent observable.
The order parameter
indicates the (time dependent) average orientation of the molecule w.r.t. ~the direction of the EF.}
If the molecule is perfectly perpendicular to the EF, then $S_\theta = 1$, if is perfectly parallel then $S_\theta = -2$, if it has no directional
preference, then $S_\theta = 0$. 
From $t=0$ to roughly 10~ps, the orientational order parameter rapidly
decreases from 0 to $-0.15$, which means a weak alignment of the molecule to the external field.
Then from $t=0$ to 500~ps, the molecule slowly changes to the orientation
that is perpendicular to the EF.
\recheck{In fact,
  {the vector of the geometric direction tends to be perpendicular
    to the dipole moment,
    % \textit{NOTE: $\frac{\pi}{2}$ corresponds to $S_\theta = 1$, however,
    %   we only have 0.4. So would should say a ``tendency of perpendicular''
    %   rather than ``perfectly perpendicular'' ($\frac{\pi}{2}$).}
  }  
  % the angle between the vector of the dipole moment and
  % the vector of the geometric direction is always close to
  % $\frac{\pi}{2}$;
  if one observes the correlation between the value
  of the dipole moment (which tends to be more parallel to the EF as
  its value increases) and the angle of the vector of the geometric
  direction, the result above can be easily explained.}
This orientational change is linked to the observed slowest
time scale of the molecular conformational change.
